Computing Entropies with Nested Sampling

### 2017-12-08 talk/ms-abridged.tex
Attempt at shorter talk

\usetheme{Warsaw}
\usecolortheme{crane}

@@

\title{Computing Entropies with Nested Sampling}
\author{Brendon J. Brewer}
\institute{Department of Statistics\\
The University of Auckland}
\date{\color{blue}\url{https://www.stat.auckland.ac.nz/~brewer/}}

@@

% New slide
\begin{frame}
\frametitle{What is entropy?}

Firstly, I'm talking about information theory, not thermodynamics (though the
two are connected).

@@

% New slide
\begin{frame}
\frametitle{Information Theory}

The fundamental theorem of information theory

% New slide
\begin{frame}
\frametitle{Information Theory}

The fundamental theorem of information theory
\begin{block}{Theorem}
If you take the log of a probability, it seems like you understand profound
truths.
\end{block}

@@

It is a real-valued property of the distribution.

@@

% New slide
\begin{frame}
\frametitle{Relative entropy}

Consider two discrete probability distributions with probabilities
$\boldsymbol{p} = \{p_i\}$ and $\boldsymbol{q} = \{q_i\}$.
The relative entropy is
\begin{align}
H(\boldsymbol{p}; \boldsymbol{q}) &= -\sum_i p_i \log\left(\frac{p_i}{q_i}\right)
\end{align}
Without the minus sign,
it's the `Kullback-Leibler divergence', and is more fundamental than the
Shannon entropy. With uniform $\boldsymbol{q}$, it reduces to the Shannon
entropy (up to an additive constant).
\end{frame}

@@

Heuristic: standard deviation quantifies uncertainty `horizontally',
entropy does it `vertically'.

@@

This generalises log-volume, as defined {\bf with respect to} $dx$.

\begin{alertblock}{Important}
Differential entropy is {\em coordinate-system dependent}.
\end{alertblock}

>>>>

\usetheme{Copenhagen}
\usecolortheme{crane}

@@

\title{Computing Entropies with Nested Sampling}
\author{Brendon J. Brewer}
\institute{Department of Statistics\\
The University of Auckland}
\date{\color{blue}\bf \url{https://www.stat.auckland.ac.nz/~brewer/}}

@@

% New slide
\begin{frame}
\frametitle{The fundamental theorem of information theory}

@@

% New slide
\begin{frame}
\frametitle{The fundamental theorem of information theory}

\begin{block}{Theorem}
If you take the log of a probability, it seems like you understand profound
truths.
\end{block}

@@

It is a real-valued property of the distribution.
There is also relative entropy / KL divergence,
which is a function of two distributions, and is more fundamental.

@@

@@

@@

This generalises log-volume, as defined {\bf with respect to} $dx$.
\pause
\begin{alertblock}{Important}
Differential entropy is {\em coordinate-system dependent}.
\end{alertblock}
